### 2015-05-12 evaluation.tex
Some comments on the experimental section.  Though I realised I commented on the section you told me was not yet ready - oops.  Jeroen would hate that.

We performed experiments to validate the usefulness of the \whoop approach (see Section~\ref{whoop}). We first present race-checking results from running \whoop and \corral on \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel distribution\footnote{\url{https://www.kernel.org}}. We then evaluate the runtime performance and scalability of \corral and \whoop + \corral with different yield instrumentation granularities and context-switch bounds, to show that \whoop can efficiently accelerate bug-finding with \corral.

\noindent
\textbf{Experimental Setup}\xspace\xspace We performed all experiments on a 3.40GHz Intel Core i7-2600 CPU with 16GB RAM running Ubuntu Linux 12.04.5 LTS, LLVM 3.5, SMACK 1.5, Z3 4.3.2, Boogie rev. 4192 and \corral rev. 534.

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose non-trivial drivers from many different domains: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We had to manually model the environment for these drivers, a process that required two months of work.

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of racy pairs that \whoop identified (\#Racy Pairs); number of racy memory regions that \whoop reported (\#Racy MRs); and number of races that \corral discovered (\#Races Found) using a context-switch bound of 2. We did not discover new races with larger context-switch bounds. This might mean that these drivers only have races that manifest with at least one or two context-switches, or that \corral hit its bounds before discovering a deeper race.

We can see that \whoop reports more races than \corral. This is expected, as our approach uses over-approximation (to find all races), and thus can introduce false bugs. On the other hand, \corral is precise, but can miss races as it explores programs under pre-defined bounds. Another issue with \corral is loop coverage due to unsound loop unrolling. To tackle this, we enable the built-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

Most of the races that \whoop and \corral discovered can be classified in two cases. The first case is accessing a global counter (or flag) from concurrent entry points, without a lock. This might be for performance, and indeed a lot of the races we found might be benign. Even benign races, though, can lead to undefined behavior according to the C standard, which can cause many issues (e.g. with compilers). The second case is an entry point accessing a field of an object (either global or passed as a parameter) without a lock. This can lead to a race if another entry point simultaneously accesses the same field of the same object.

@@

\noindent
\textbf{Granularity of Context-Switches}\xspace\xspace Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks. All reported times are in seconds and averages of three runs. \corral was given a time budget of 10 hours (T.O. denotes a tool timeout) and a context-switch bound (csb) of 2, 5 and 9. By default, \corral instruments context-switches (i.e. \texttt{yield} statements) in all visible operations (denoted by \yieldall). \whoop + \corral, instead, uses two different context-switch instrumentation granularities, \yieldcoarse and \yieldmr (see Section~\ref{whoop:bugfinding}), and also uses all the optimisations discussed in Section~\ref{whoop:optimizations}.

>>>>

We performed experiments to validate the usefulness of the \whoop approach (Section~\ref{whoop}) and its combination with \corral (Section~\ref{corral}). We first present race-checking results from running \whoop and \corral on \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel distribution.\footnote{\url{https://www.kernel.org}} We then evaluate the runtime performance and scalability of \corral and \whoop + \corral with different yield instrumentation granularities and context-switch bounds.
Our results demonstrate that \whoop can efficiently accelerate bug-finding with \corral.

\noindent
\textbf{Experimental Setup}\xspace\xspace We performed all experiments on a 3.40GHz Intel Core i7-2600 CPU with 16GB RAM running Ubuntu Linux 12.04.5 LTS, LLVM 3.5, SMACK 1.5, Z3 4.3.2, Boogie rev. 4192 and \corral rev. 534.  \ADComment{Perhaps also state that Boogie runs inside Mono, and give the Mono version?}

\noindent
\textbf{Benchmarks}\xspace\xspace We evaluate our methodology against \sizeOfBenchmarks drivers taken from the 4.0 Linux kernel. We chose non-trivial drivers from several domains: block; char; ethernet; near field communication (nfc); universal serial bus (usb); and watchdog. We had to manually model the environment for these drivers, a process that required approximately two months of work.  \ADComment{Question from me: are block and char domains, in the same way that ethernet is a domain?}

\ADComment{At this point, is it absolutely clear what we do with \corral?  I.e., that we consider each entry point pair separately, and do not employ any abstraction to model additional threads?}

\noindent
\textbf{Race-Checking}\xspace\xspace Table~\ref{tab:stats} presents statistics for all our benchmarks: lines of code (LoC); number of entry point pairs (\#Pairs); number of SMACK memory regions (\#MRs); number of potentially racy pairs identified by \whoop (\#Racy Pairs); number of potentially racy memory regions reported by \whoop (\#Racy MRs); and number of races that discovered by \corral using a context-switch bound of 2 (\#Races Found). We did not discover new races with larger context-switch bounds. This might mean that these drivers only have races that manifest with at least one or two context-switches, or that \corral hit its bounds before discovering a deeper race.

We can see that \whoop reports more races than \corral. This is expected, since \whoop employs an over-approximating shared state abstraction to conservatively model the effects of additional threads when analysing an entry point pair, and because lockset analysis is inherently imprecise; both factors can lead to false alarm race reports.  On the other hand, \corral is precise, but can miss races because only a limited number of context switches are considered.  Another issue with \corral is loop coverage due to unsound loop unrolling. To tackle this, we enable the built-in loop over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.  \ADComment{Would this be a good point to reiterate the fact that when we apply \corral to an entry point pair, we are just checking the specific pair of threads, and are not accounting for actions by other threads?}

Most of the races that \whoop and \corral discovered can be classified in two cases. The first case is accessing a global counter (or flag) from concurrent entry points, without a lock. This might be for performance, and indeed a lot of the races we found might be benign. Even benign races, though, lead to undefined behavior according to the C standard, and it is well known that undefined behaviors can lead to unexpected results when combined with aggressive compiler optimizations. The second case is an entry point accessing a field of an object (either global or passed as a parameter) without a lock. This can lead to a race if another entry point simultaneously accesses the same field of the same object.

@@

\noindent
\textbf{Granularity of Context-Switches}\xspace\xspace Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks. All reported times, in seconds, are averages over three runs. \ADComment{Can we say something about variance?  Perhaps ask Nathan what best to say?} \corral was given a time budget of 10 hours (T.O. denotes a tool timeout), and we considered context-switch bounds (csb) of 2, 5 and 9. By default, \corral instruments context-switches (i.e. \texttt{yield} statements) in \ADComment{instead of ``in'' can we say either ``before'' or ``after'' (not sure which is accurate)} all visible operations (denoted \yieldall). \whoop + \corral, instead, uses two different context-switch instrumentation granularities, \yieldcoarse and \yieldmr (see Section~\ref{whoop:bugfinding}), and also uses all the optimisations discussed in Section~\ref{whoop:optimizations}.
